---
title: "Effect of the number of SLAM filter particles parameter on the average RAM utilisation"
output:
  html_notebook: default
---

We need to analyse the impact of two possible values for the number of SLAM filter particles on the average RAM utilisation. We have chosen two values for the number of particles, namely, `5`, which is used in the primary experiments, and `30`, which is the default value in Gmapping ROS package. First, let us load the necessary libraries:
```{r}
if (!require('tidyverse')) install.packages('tidyverse', dependencies = TRUE); library('tidyverse')
if (!require('ggpubr')) install.packages('ggpubr', dependencies = TRUE); library('ggpubr')
if (!require('rstatix')) install.packages('rstatix', dependencies = TRUE); library('rstatix')
if (!require('ggplot2')) install.packages('ggplot2', dependencies = TRUE); library('ggplot2')
```

Next, we load the raw data output of the performed experiment:
```{r}
data <- read.csv(file = '../../raw_data/test_particles/run_table.csv')
data$avg_memory_util <- data$avg_memory_util / 1000000
data <- subset(data, select = c(X__run_id, particles, avg_memory_util))
data$particles <- as.factor(data$particles)
data$avg_memory_util <- as.numeric(data$avg_memory_util)
```

Mean and standard deviation of the average RAM utilisation for both frame particle number values are displayed in the summary below:
```{r}
data %>%
  group_by(particles) %>%
  get_summary_stats(avg_memory_util, type = "mean_sd")
```

We can visualize the distribution of the data for both frame particle number treatments via boxplot below:
```{r}
#pdf(file = "../../figures/particles_effect/ram-utilisation.pdf")

ggplot(data, aes(x = particles, y = avg_memory_util, fill = particles)) +
  theme_bw() + 
  geom_boxplot(width = 0.9, outlier.size = 0.5) + 
  stat_summary(fun = mean, color = 'black', geom = 'point', shape = 5, size = 1.5) +
  labs(x = "Number of SLAM filter particles (#)", y = "Average RAM utilisation (MB)") +
  theme(legend.position = 'none') +
  theme(
    axis.title.x = element_text(size = 12, family = 'sans'),
    axis.title.y = element_text(size = 12, family = 'sans'),
    axis.text.x = element_text(size = 11, family = 'sans'),
    axis.text.y = element_text(size = 11, family = 'sans')
  )

#dev.off()
```

# Welch's t-test

We will apply the Welch's t-test to compare the means of both treatments. This is the default t-test in R and the safer alternative to standard Student's t-test, when we cannot assume that variances of both treatments are equal. However, we still need to check if other assumptions for applying this test are met. The first assumption of having independent observations, i.e. each subject belongs to only one group and there are no relationships between observations within the two groups, is satisfied inherently by the very way in which the experiment is conducted.

## No significant outliers

Next assumption that we need to check is if there are no significant outliers in neither of the two groups:
```{r}
data %>% 
  group_by(particles) %>%
  identify_outliers(avg_memory_util)
```

As the result above shows, there are two outliers, for runs 11 and 14, respectfully, while run 14 is considered as extreme. Since the assumption of no significant outliers is not satisfied, we will proceed with non-parametric alternative to independent two samples t-test, which is Wilcoxon rank sum test.

# Wilcoxon rank-sum test

We can execute Wilcoxon rank-sum test without any assumptions checking:

```{r}
data %>%
  rstatix::wilcox_test(avg_memory_util ~ particles) %>%
  add_significance()
```
The resulting p-value of $1.08 \cdot 10^{-5}$ is lower than `0.05` reference value for the `0.95` confidence interval. This means that the obtained p-value is statistically significant and we can reject $H_0$, which states that the means of the average RAM utilisation for both particle number values are equal. We can hereby conclude that the number of filter particles parameter has a statistically significant effect on the average RAM utilisation.

## Effect size

We establish with the Wilcoxon rank sum test results that the particles numbers parameter has a statistically significant effect on the average RAM utilisation Now we need estimate the magnitude of the said effect with Wilcoxon effect size test based on the Z statistics:
```{r}
data %>%
  wilcox_effsize(avg_memory_util ~ particles)
```
The resulting effect size of `R = 0.8451543` is interpreted as _large_^[https://www.researchgate.net/publication/303919832_The_need_to_report_effect_size_estimates_revisited_An_overview_of_some_recommended_measures_of_effect_size]. This is indeed clear from the distribution of data for both treatments in the boxplot above, where we can see that the average RAM utilisation mean for `5` particles is `877374634 B`, while it is `881817482 B` for `30` particles. 
