---
title: "Unknown map experiment - CPU usage"
output:
  html_notebook: default
---

We are analyzing if offloading any of the three functionalities (i.e., SLAM, navigation and object recognition) has a statistically significant effect on the average CPU usage. Fist, install and load the required libraries:
```{r}
if (!require('tidyverse')) install.packages('tidyverse', dependencies = TRUE); library('tidyverse')
if (!require('ggpubr')) install.packages('ggpubr', dependencies = TRUE); library('ggpubr')
if (!require('rstatix')) install.packages('rstatix', dependencies = TRUE); library('rstatix')
if (!require('ggplot2')) install.packages('ggplot2', dependencies = TRUE); library('ggplot2')
if (!require('lmPerm')) install.packages('lmPerm', dependencies = TRUE); library('lmPerm')
if (!require('emmeans')) install.packages('emmeans', dependencies = TRUE); library('emmeans')
```

Next, load the experiment results from the csv file:
```{r}
data <- read.csv(file = '../../raw_data/unknown_map_experiment/run_table.csv')
data <- subset(data, select = c(X__run_id, slam_offloaded, navigation_offloaded, obj_recognition_offloaded, avg_cpu_util))
data$slam_offloaded <- as.factor(data$slam_offloaded)
data$navigation_offloaded <- as.factor(data$navigation_offloaded)
data$obj_recognition_offloaded <- as.factor(data$obj_recognition_offloaded)
data$avg_cpu_util <- as.numeric(data$avg_cpu_util)
```

Mean and standard deviation of the average CPU usage for each treatment are displayed in the summary below:
```{r}
data %>%
  group_by(slam_offloaded, navigation_offloaded, obj_recognition_offloaded) %>%
  get_summary_stats(avg_cpu_util, type = "mean_sd")
```

# Three-Way ANOVA test

Since there are three independent variables in the experiment, we opt for three-way ANOVA test. First, we need to check if all assumptions for performing the test are met. The fist assumption of having independent observations is met inherently according to the way the experiment itself is conducted. 

## No extreme outliers assumption

Next, we need to check if there are any extreme outliers:

```{r}
data %>%
  group_by(slam_offloaded, navigation_offloaded, obj_recognition_offloaded) %>%
  identify_outliers(avg_cpu_util)
```

The results show that there are seven outliers, in runs 14, 66, 11, 42, 37, 54 and 57, respectfully, while run 42 represents an extreme outlier. We should remove the said run from the dataset, but we can also leave it if we believe that it will not have a significant effect on the obtained results. Let us make an alternative dataset with the extreme outlier being removed:

```{r}
data_without_outliers <- data
data_without_outliers <- data_without_outliers[!(data_without_outliers$X__run_id == 'run_42'),]
```
We proceed to check the next assumption - normality of the data. 

## Normality assumption by analyzing the model residuals

The first way to check normality assumption is by analyzing the model residuals:

```{r}
model  <- lm(avg_cpu_util ~ slam_offloaded*navigation_offloaded*obj_recognition_offloaded, data = data)
ggqqplot(residuals(model))
```

The QQ-plot in the figure above shows that most of the residuals are approximately alongside the reference line, with some exceptions around the edges. We will complement the QQ-plot with Shapiro-Wilk test to conclude if normality is indeed satisfied:
```{r}
shapiro_test(residuals(model)) %>%
  add_significance()
```
The p-value of `0.175` shows that we cannot formally reject $H_0$, which states that the data is normally distributed. We can conclude that the data is indeed normally distributed.

## Normality assumption by each group

The other way to check normality is by analyzing if samples for each possible treatment combination are normally distributed:

```{r}
data %>%
  group_by(slam_offloaded, navigation_offloaded, obj_recognition_offloaded) %>%
  shapiro_test(avg_cpu_util) %>%
  add_significance()
```

Shapiro-Wilk test per each group yields p-value higher than `0.05` for all groups, except for the two groups where SLAM and object recognition are not offloaded but navigation is (`p = 0.035`) and when SLAM is offloaded but navigation and object recognition are not. If we execute the test on the dataset without extreme outliers:
```{r}
data_without_outliers %>%
  group_by(slam_offloaded, navigation_offloaded, obj_recognition_offloaded) %>%
  shapiro_test(avg_cpu_util) %>%
  add_significance()
```
we still have one group that is not normally distributed. Nevertheless, we will proceed with checking further assumptions of three-way ANOVA test since normality of residuals is satisfied. Checking normality per group yields more accurate results when there are many data points per group, which is not the case in this experiment (10 data points per group). The QQ-plots per group are depicted below, where most of the data points seem to be aligned with the reference line:

```{r}
ggqqplot(data, "avg_cpu_util", ggtheme = theme_bw()) +
  facet_grid(slam_offloaded + navigation_offloaded ~ obj_recognition_offloaded, labeller = "label_both")
```

## Homogeneity of variance assumption

Homogeneity of variance assumption can be checked via Levene’s test:

```{r}
data %>% 
  levene_test(avg_cpu_util ~ slam_offloaded*navigation_offloaded*obj_recognition_offloaded) %>%
  add_significance()
```
Since the p-value is lower than `0.05`, the result of the Levene's test is statistically significant, hence the null hypothesis stating that homogeneity of variance holds is rejected. If we compute the Levene's test on dataset without extreme outliers:
```{r}
data_without_outliers %>% 
  levene_test(avg_cpu_util ~ slam_offloaded*navigation_offloaded*obj_recognition_offloaded) %>%
  add_significance()
```
we can draw the same conclusion - homogeneity of variance assumption is not satisfied hence we cannot apply three-way ANOVA. This is why we turn to its non-parametric alternative - permutation test.

# Permutation test

Execution of permutation test yields the following result:
```{r}
summary(aovp(avg_cpu_util ~ slam_offloaded*navigation_offloaded*obj_recognition_offloaded, data = data, perm="Prob", maxIter=500000, nCycle=100, Ca=0.001))
```
Test results show that the main effects are significant for all three independent variables, but there is also a significant two-way interaction between SLAM and object recognition. If we compare these results with three-way ANOVA results, that is considered as _robust_ when sample sizes are equal for each group, which is indeed the case in this experiment:
```{r}
data %>% 
  anova_test(avg_cpu_util ~ slam_offloaded*navigation_offloaded*obj_recognition_offloaded)

```
we see that the obtained p-values and conclusions are very similar to those drawn from the permutation test. The eta-squared effect sizes are:
```{r}
res.aov <- aov(avg_cpu_util ~ slam_offloaded*navigation_offloaded*obj_recognition_offloaded, data = data)
eta_squared(res.aov)
```

## Pairwise comparisons for significant slam_offloaded:obj_recognition_offloaded two-way interaction 

Since we have a significant two-way interaction between SLAM and object recognition, we need to perform multiple pairwise comparison to determine which group means are different. We will use estimated marginal means test for this purpose with Bonferroni adjustment. We will group the result by object recognition, since the main effect of object recognition is greater than the one of SLAM, according to three-way ANOVA results:
```{r}
data %>% 
  group_by(obj_recognition_offloaded) %>%
  emmeans_test(avg_cpu_util ~ slam_offloaded, p.adjust.method = "bonferroni", detailed = TRUE) 
```
The pairwise comparison between offloading and not offloading SLAM groups when object recognition is either offloaded or not are all significant.

# Results

In the box plot below, we can indeed see that when object recognition is both offloaded and not, the average CPU usage is lower when SLAM is offloaded. However, if we compare the means of two groups on the left and right, when object recognition is not offloaded and offloaded, we can see that offloading object recognition does have an enormous overall effect on the average CPU usage:
```{r}
#pdf(file = "../../figures/unknown_map_experiment/cpu-usage-1.pdf")

ggplot(data, aes(x = obj_recognition_offloaded, y = avg_cpu_util, fill = slam_offloaded)) +
  theme_bw() + 
  geom_boxplot(width = 0.9, outlier.size = 0.5) + 
  stat_summary(fun = mean, color = 'black', geom = 'point', shape = 5) +
  labs(x = "Object recognition offloaded", y = "Average CPU usage (%)") +
  guides(fill=guide_legend(title="SLAM offloaded")) +
  theme(
    axis.title.x = element_text(size = 12, family = 'sans'),
    axis.title.y = element_text(size = 12, family = 'sans'),
    axis.text.x = element_text(size = 11, family = 'sans'),
    axis.text.y = element_text(size = 11, family = 'sans'),
    legend.title = element_text(size = 12, family = 'sans'),
    legend.text = element_text(size = 11, family = 'sans'),
    legend.position = "top"
  )

#dev.off()
```

Both two-way anova and permutation test yielded the significant main effect when navigation is offloaded, without interactions with the other two factors:
```{r}
#pdf(file = "../../figures/unknown_map_experiment/cpu-usage-2.pdf")

ggplot(data, aes(x = navigation_offloaded, y = avg_cpu_util, fill = navigation_offloaded)) +
  theme_bw() + 
  geom_boxplot(width = 0.9, outlier.size = 0.5) + 
  stat_summary(fun = mean, color = 'black', geom = 'point', shape = 5) +
  labs(x = "Navigation offloaded", y = "Average CPU usage (%)") +
  theme(legend.position = "none") +
  theme(
    axis.title.x = element_text(size = 12, family = 'sans'),
    axis.title.y = element_text(size = 12, family = 'sans'),
    axis.text.x = element_text(size = 11, family = 'sans'),
    axis.text.y = element_text(size = 11, family = 'sans'),
    
  )

#dev.off()
```
Even though CPU usage is lower when navigation is offloaded, the boxplot above shows that the improvement is not that significant. 
