---
title: "Effect of camera frame rate parameter on the average feature extraction time"
output:
  html_notebook: default
---

We need to analyse the impact of two possible values of camera frame rate on the average feature extraction time. We have chosen two values for frame rate, namely, `20fps`, which is used in the primary experiments, and three times higher frame rate, `60fps`. First, let us load the necessary libraries:
```{r}
if (!require('tidyverse')) install.packages('tidyverse', dependencies = TRUE); library('tidyverse')
if (!require('ggpubr')) install.packages('ggpubr', dependencies = TRUE); library('ggpubr')
if (!require('rstatix')) install.packages('rstatix', dependencies = TRUE); library('rstatix')
if (!require('ggplot2')) install.packages('ggplot2', dependencies = TRUE); library('ggplot2')
```

Next, we load the raw data output of the performed experiment:
```{r}
data <- read.csv(file = '../../raw_data/test_frame_rate/run_table.csv')
data <- subset(data, select = c(X__run_id, frame_rate, avg_extraction_time_ms))
data$frame_rate <- as.factor(data$frame_rate)
data$avg_extraction_time_ms <- as.numeric(data$avg_extraction_time_ms)
```

Mean and standard deviation of the average feature extraction time for both frame rates are displayed in the summary below:
```{r}
data %>%
  group_by(frame_rate) %>%
  get_summary_stats(avg_extraction_time_ms, type = "mean_sd")
```

We can visualize the distribution of the data for both frame rate treatments via boxplot below:
```{r}
#pdf(file = "../../figures/frame_rate_effect/extraction-time.pdf")

ggplot(data, aes(x = frame_rate, y = avg_extraction_time_ms, fill = frame_rate)) +
  theme_bw() + 
  geom_boxplot(width = 0.9, outlier.size = 0.5) + 
  stat_summary(fun = mean, color = 'black', geom = 'point', shape = 5, size = 1.5) +
  labs(x = "Frame rate (fps)", y = "Average feature\nextraction time (ms)") +
  theme(legend.position = 'none') +
  theme(
    axis.title.x = element_text(size = 12, family = 'sans'),
    axis.title.y = element_text(size = 12, family = 'sans'),
    axis.text.x = element_text(size = 11, family = 'sans'),
    axis.text.y = element_text(size = 11, family = 'sans')
  )

#dev.off()
```

# Welch's t-test

We will apply the Welch's t-test to compare the means of both treatments. This is the default t-test in R and the safer alternative to standard Student's t-test, when we cannot assume that variances of both treatments are equal. However, we still need to check if other assumptions for applying this test are met. The first assumption of having independent observations, i.e. each subject belongs to only one group and there are no relationships between observations within the two groups, is satisfied inherently by the very way in which the experiment is conducted.

## No significant outliers

Next assumption that we need to check is if there are no significant outliers in neither of the two groups:
```{r}
data %>% 
  group_by(frame_rate) %>%
  identify_outliers(avg_extraction_time_ms)
```

As the result above shows, there are three outliers, for runs 14, 4 and 19, respectfully, but neither is considered extreme in terms of statistical significance. We can proceed with other assumptions.

## Normality

Next, we need to check if data in both groups stem from a normal distribution. To do so, we first need to draw QQ-plot for both groups:
```{r}
ggqqplot(data, x = "avg_extraction_time_ms", facet.by = "frame_rate")
```
Data points on both figures seem to be more or less aligned alongside the respective reference lines, with a few excpetional data points, hence we may assume that normality assumption is satisfied for both frame rate treatment groups. Nevertheless, we need to perform Shapiro-Wilk test for each group to confirm the assumptions drawn from the QQ-plots:
```{r}
data %>%
  group_by(frame_rate) %>%
  shapiro_test(avg_extraction_time_ms)
```
As the results of the test show, p-value for frame rate `60fps` group is higher than `0.05` (`p = 0.865`), but it less than `0.05` for `20fps` treatment group (`p = 0.0059`). This means that the data in the latter group is not normal distributed since the null hypothesis in Shapiro's test, stating that data does stem from a normal distribution, is rejected. Since the normality assumption for performing Welch's t-test is not satisfy, we need to conduct its non-parametric alternative, i.e. Wilcoxon rank-sum test.

# Wilcoxon rank-sum test

We can execute Wilcoxon rank-sum test without any assumptions checking:

```{r}
data %>%
  rstatix::wilcox_test(avg_extraction_time_ms ~ frame_rate) %>%
  add_significance()
```
The resulting p-value of $0.0232$ is lower than `0.05` reference value for the `0.95` confidence interval. This means that the obtained p-value is statistically significant and we can reject $H_0$, which states that the means of the average feature extraction time for both frame rate values are equal. We can hereby conclude that camera frame rate parameter has a statistically significant effect on the average feature extraction time.

## Effect size

We establish with the Wilcoxon rank sum test results that the frame rate parameter has a statistically significant effect on the average feature extraction time. Now we need estimate the magnitude of the said effect with Wilcoxon effect size test based on the Z statistics:

```{r}
data %>%
  wilcox_effsize(avg_extraction_time_ms ~ frame_rate)
```
The resulting effect size of `R = 0.5070926` is interpreted as _large_^[https://www.researchgate.net/publication/303919832_The_need_to_report_effect_size_estimates_revisited_An_overview_of_some_recommended_measures_of_effect_size]. This result is not as straightforward to conclude from the distribution of data for both treatments in the boxplot above, where we can see that the average feature ectraction time mean for `20fps` frame rate is `5.445 ms`, while it is `5.318 ms` for `60fps`. 
