---
title: "Effect of camera frame rate parameter on the total mission execution time"
output:
  html_notebook: default
---

We need to analyse the impact of two possible values of camera frame rate on the total mission execution time. We have chosen two values for frame rate, namely, `20fps`, which is used in the primary experiments, and three times higher frame rate, `60fps`. First, let us load the necessary libraries:
```{r}
if (!require('tidyverse')) install.packages('tidyverse', dependencies = TRUE); library('tidyverse')
if (!require('ggpubr')) install.packages('ggpubr', dependencies = TRUE); library('ggpubr')
if (!require('rstatix')) install.packages('rstatix', dependencies = TRUE); library('rstatix')
if (!require('ggplot2')) install.packages('ggplot2', dependencies = TRUE); library('ggplot2')
```

Next, we load the raw data output of the performed experiment:
```{r}
data <- read.csv(file = '../../raw_data/test_frame_rate/run_table.csv')
data <- subset(data, select = c(X__run_id, frame_rate, mission_execution_s))
data$frame_rate <- as.factor(data$frame_rate)
data$mission_execution_s <- as.numeric(data$mission_execution_s)
```

Mean and standard deviation of the total mission execution time for both frame rates are displayed in the summary below:
```{r}
data %>%
  group_by(frame_rate) %>%
  get_summary_stats(mission_execution_s, type = "mean_sd")
```

We can visualize the distribution of the data for both frame rate treatments via boxplot below:
```{r}
ggplot(data, aes(x = frame_rate, y = mission_execution_s, fill = frame_rate)) +
  theme_bw() + 
  geom_boxplot(width = 0.9, outlier.size = 0.5) + 
  stat_summary(fun = mean, color = 'black', geom = 'point', shape = 5, size = 1.5) +
  labs(x = "Frame rate (fps)", y = "Total mission execution time (s)") +
  theme(legend.position = 'none') +
  theme(
    axis.title.x = element_text(size = 12, family = 'sans'),
    axis.title.y = element_text(size = 12, family = 'sans'),
    axis.text.x = element_text(size = 11, family = 'sans'),
    axis.text.y = element_text(size = 11, family = 'sans'),
    
  )
```

# Welch's t-test

We will apply the Welch's t-test to compare the means of both treatments. This is the default t-test in R and the safer alternative to standard Student's t-test, when we cannot assume that variances of both treatments are equal. However, we still need to check if other assumptions for applying this test are met. The first assumption of having independent observations, i.e. each subject belongs to only one group and there are no relationships between observations within the two groups, is satisfied inherently by the very way in which the experiment is conducted.

## No significant outliers

Next assumption that we need to check is if there are no significant outliers in neither of the two groups:
```{r}
data %>% 
  group_by(frame_rate) %>%
  identify_outliers(mission_execution_s)
```

As the result above shows, there are two outliers, for runs 2 and 18, respectfully, while run 18 is considered as an extreme outlier. Since the assumption of no significant outliers is not satisfied, we will proceed with non-parametric alternative to independent two samples t-test, which is Wilcoxon rank sum test.

# Wilcoxon rank-sum test

We can execute Wilcoxon rank-sum test without any assumptions checking:

```{r}
data %>%
  rstatix::wilcox_test(mission_execution_s ~ frame_rate) %>%
  add_significance()
```
The resulting p-value of $0.0753$ is higher than `0.05` reference value for the `0.95` confidence interval. This means that the obtained p-value is not statistically significant and we cannot reject $H_0$, which states that the means of the total mission execution time for both frame rate values are equal. We can hereby conclude that the frame rate parameter does not have a statistically significant effect on the total mission execution time.
